---
title: "RProject"
author: "Piotr Jówko"
output: github_document
date: "`r format(Sys.time(), '%d %B, %Y')`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Table of contents
1. [Summary](#summary)
2. [Used libraries](#libriaries)
3. [Data loading](#dataLoading)
4. [Data Cleaning](#cleaning)
5. [Basic statistics](#baseStats)
6. [Correlations](#correlations)

## <a name="summary"></a> Summary
//TODO

## <a name="libriaries"></a> Used libraries
- dplyr
- tidyr
- ggplot2
- corrplot
- caret
```{r libriaries, include=FALSE}
library(dplyr);
library(tidyr)
library(ggplot2);
library(corrplot)
library(caret);

set.seed(23);
```


## <a name="dataLoading"></a> Data loading

Reading data from file:

```{r dataLoading, cache=TRUE}
panels <- read.csv(file = 'solar_panels.csv')
#panels <- head(panels, 1000) #used for testing only
```

## <a name="cleaning"></a> Data Cleaning
All rows with any NA values were removed. </br>
Also columns: id, data, anno were removed. Id is not needed in analysis. </br>
Data is redundant, because date is stored in others columns and there it is normalized(ora, day, anno). </br>
Anno had only two values: 2012 and 2013, with will not be needed in analysis.
```{r cleaning}
panels <- panels %>% drop_na() %>% select(-id, -anno, -data)
```

Also some fields were renamed, because they names were confusing or unclear. 
```{r cleaning2}
panels <- panels %>% rename(day_365 = day, latitude = lat, longitude = lon, hour = ora, room_temperature = temperatura_ambiente, irradiation = irradiamento)
interactiveChart <- select(panels, latitude, longitude, hour, kwh)
```

## <a name="baseStats"></a> Basic statistics
Number of rows and columns:
```{r baseStats, echo=FALSE}
dim(panels)
```

Data summary:
```{r baseStats2, echo=FALSE}
summary(panels)
```

## <a name="correlations"></a> Correlations
In thi step correlation matrix is calculated(correlation with each pair of columns). <br>
For clarity some rows(with no correlation greater than 0.4 or less than -0.4) were removed.
```{r correlations, fig.height = 10, fig.width = 10}
corMatrix <- cor(panels)

# Change diagonal values from 1 to 0. Needed for weak correlations removal.
for(i in 1:dim(corMatrix)[1]) {
    corMatrix[i, i] <- 0;
}

# Remove "weak" correlations.
strongCor <- corMatrix[apply(corMatrix, MARGIN = 1, function(x) any(x > 0.40 | x < -0.40)), ]

corrplot(strongCor, method = "square", type="lower", diag=FALSE)
```

As it seen on chart, kwh have strong positive correlation with irriadiation and strong negative correlation with humidity with is obvious. 
